---
title: "DW"
author: "Huy Nguyen"
date: "2022-07-20"
output: html_document
---

Load Data
```{r}
library(readxl)
library(tidyverse)
library(psych)
library(survival)
library(coin)

dat = readRDS("dat.rds")
```

DW Dataframe
```{r}
dat = dat[, c("ID", "Sex", "Condition", "DW_Latency", "DW_LatencyD2", 
             "DW_tHide", "DW_tHideD2", "DW_nExit", "DW_nExitD2")]
```

### **ANALYSIS**
#### Filter Data
```{r} 
test = dat %>% group_by(ID) %>%
  summarise(
    Sex, 
    Condition, 
    Status = paste(Sex, Condition, sep = "_"),
    Latency = mean(c(DW_Latency, DW_LatencyD2)),
    tHide = mean(c(DW_tHide, DW_tHideD2)),
    nExit = mean(c(DW_nExit, DW_nExitD2))
  ) 
test$Status = as.factor(test$Status)

saveRDS(test, "DW.rds")
```

#### Mean, Median, & SD
##### Latency 
```{r}
tapply(test$Latency, test$Status, mean)
tapply(test$Latency, test$Status, median)
tapply(test$Latency, test$Status, sd)
```
Only slight difference between mean & median. Owl groups & females take longer to come out but only in mere seconds. 

##### Hide
```{r}
tapply(test$tHide, test$Status, mean)
tapply(test$tHide, test$Status, median)
tapply(test$tHide, test$Status, sd)
```
Some noticeable differences between mean (1st row) & median (2nd row); especially females. However, no skew was found. Opposite trend for male & female with with control spend more time hiding for female and vice versa. 

##### Exit
```{r}
tapply(test$nExit, test$Status, mean)
tapply(test$nExit, test$Status, median)
tapply(test$nExit, test$Status, sd)
```
Very similar results, differences are within 1 second so I don't expect to see significant results here. 

#### Skew
```{r}
tapply(test$Latency, test$Status, skew) #Latency
tapply(test$tHide, test$Status, skew) #Hide
tapply(test$nExit, test$Status, skew) #Exit
```
No skew

#### Test for Equal Variance
```{r}
bartlett.test(Latency ~ Status, data = test) #p-value = 0.0472
bartlett.test(tHide ~ Status, data = test) #p-value = 0.3776
bartlett.test(nExit ~ Status, data = test) #p-value = 0.6562
```
Only Leave Latency violated the normality assumption. Multiple ANOVA tests were ran just to be sure. 

#### ANOVA
```{r}
ANOVA = list()
```

```{r}
#Leave Latency
ANOVA$LeaveLatency = lm(Latency ~ Status, data = test) #F-statistic: 3.177 on 3 and 15 DF,  p-value: 0.05487
summary(ANOVA$LeaveLatency)

kruskal.test(Latency ~ Status, data = test) #p-value = 0.03678

oneway.test(Latency ~ Status, data = test, var.equal = FALSE) #p-value =0.0275
```
Leave Latency differences are significant across groups

```{r}
#Hide
ANOVA$Hide = lm(tHide ~ Status, data = test) #p-value: 0.1366
summary(ANOVA$Hide)

kruskal.test(tHide ~ Status, data = test) #p-value = 0.1411

oneway.test(tHide ~ Status, data = test, var.equal = FALSE) #p-value = 0.1195
```
Hide time differences are insignificant across groups 

```{r}
#Exit
ANOVA$Exit = lm(nExit ~ Status, data = test) #p-value: 0.9474
summary(ANOVA$Exit)

kruskal.test(nExit ~ Status, data = test) #p-value = 0.9731

oneway.test(nExit ~ Status, data = test, var.equal = FALSE) #p-value = 0.9049
```
Number of exits differences are insignificant across groups 

####Tukey HSD
```{r}
#Leave Latency
ANOVA$LL_aov = aov(Latency ~ Status, data = test)
TukeyHSD(ANOVA$LL_aov, conf.level = .95)
```
Only Male_control-female_owl differences are significant

```{r}
#Hide
ANOVA$H_aov = aov(tHide ~ Status, data = test)
TukeyHSD(ANOVA$H_aov, conf.level = .95)
```
No significant differences

```{r}
#Exit
ANOVA$E_aov = aov(nExit ~ Status, data = test)
TukeyHSD(ANOVA$E_aov, conf.level = .95)
```
No significant differences
