---
title: "EL"
author: "Huy Nguyen"
date: "2022-07-20"
output: html_document
---

Load Data
```{r}
library(readxl)
library(tidyverse)
library(psych)
library(survival)
library(coin)

dat = readRDS("dat.rds")
```

DW Dataframe
```{r}
dat = dat[, c("ID", "Sex", "Condition", "EL_nOpen", "EL_nClosed", 
             "EL_tOpen", "EL_tClosed", "EL_nOpenD2", "EL_nClosedD2", 
             "EL_tOpenD2", "EL_tClosedD2")]
```

### **ANALYSIS**
#### Filter Data
```{r}
test = dat %>% group_by(ID) %>%
  summarise(
    Sex, 
    Condition, 
    Status = paste(Sex, Condition, sep = "_"),
    nOpen = mean(c(EL_nOpen, EL_nOpenD2)),
    nClosed = mean(c(EL_nClosed, EL_nClosedD2)),
    tOpen = mean(c(EL_tOpen, EL_tOpenD2)),
    tClosed = mean(c(EL_tClosed, EL_tClosedD2))
  ) 
test$Status = as.factor(test$Status)

saveRDS(test, "EL.rds")
```

#### Mean, Median, & SD
##### nOpen 
```{r}
tapply(test$nOpen, test$Status, mean)
tapply(test$nOpen, test$Status, median)
tapply(test$nOpen, test$Status, sd)
```
Pretty similar number of entries across groups so I don't expect to see any significant difference. 

##### nClosed
```{r}
tapply(test$nClosed, test$Status, mean)
tapply(test$nClosed, test$Status, median)
tapply(test$nClosed, test$Status, sd)
```
Same story for closed entries. 

##### tOpen
```{r}
tapply(test$tOpen, test$Status, mean)
tapply(test$tOpen, test$Status, median)
tapply(test$tOpen, test$Status, sd)
```
Male control spent more time in open arms (consistent with previous research) while opposite for females, even then differences are only 20s. Don't see any noticeable different in time for males & females. 

#####tClosed
```{r}
tapply(test$tClosed, test$Status, mean)
tapply(test$tClosed, test$Status, median)
tapply(test$tClosed, test$Status, sd)
```
Rats spend a fuckton more time in closed arms than open arms. Similar trend within groups here. 

#### Skew
```{r}
tapply(test$nOpen, test$Status, skew) #nOpen
tapply(test$nClosed, test$Status, skew) #nClosed
tapply(test$tOpen, test$Status, skew) #tOpen
tapply(test$tClosed, test$Status, skew) #tClosed
```
No skew was found

#### Test for Equal Variance
```{r}
bartlett.test(nOpen ~ Status, data = test) #p-value = 0.9637
bartlett.test(nClosed ~ Status, data = test) #p-value = 0.932
bartlett.test(tOpen ~ Status, data = test) #p-value = 0.8951
bartlett.test(tClosed ~ Status, data = test) #p-value = 0.8559
```

#### ANOVA
```{r}
ANOVA = list()
```

```{r}
#nOpen
ANOVA$nOpen = lm(nOpen ~ Status, data = test) #p-value: 0.6373
summary(ANOVA$nOpen)
```

```{r}
#nClosed
ANOVA$nClosed = lm(nClosed ~ Status, data = test) #p-value: 0.5241
summary(ANOVA$nClosed)
```

```{r}
#tOpen
ANOVA$tOpen = lm(tOpen ~ Status, data = test) #p-value: 0.4819
summary(ANOVA$tOpen)
```

```{r}
#tClosed
ANOVA$tClosed = lm(tClosed ~ Status, data = test) #p-value: 0.5704
summary(ANOVA$tClosed)
```

#### Tukey HSD
```{r}
#nOpen
ANOVA$nOpen = aov(nOpen ~ Status, data = test)
summary(ANOVA$nOpen)
TukeyHSD(ANOVA$nOpen, conf.level = .95)
```

```{r}
#Hide
ANOVA$nClosed = aov(nClosed ~ Status, data = test)
summary(ANOVA$nClosed)
TukeyHSD(ANOVA$nClosed, conf.level = .95)
```

```{r}
#Exit
ANOVA$tOpen = aov(tOpen ~ Status, data = test)
summary(ANOVA$tOpen)
TukeyHSD(ANOVA$tOpen, conf.level = .95)
```

```{r}
#Exit
ANOVA$tClosed = aov(tClosed ~ Status, data = test)
summary(ANOVA$tClosed)
TukeyHSD(ANOVA$tClosed, conf.level = .95)
```

### Difference between tOpen & tClosed
#### Mean, Median & SD
```{r}
#Mean
sapply(select(test, tOpen, tClosed), mean) #t
sapply(select(test, nOpen, nClosed), mean) #n
```

```{r}
#Median
sapply(select(test, tOpen, tClosed), median) #t
sapply(select(test, nOpen, nClosed), median) #n
```

```{r}
#SD
sapply(select(test, tOpen, tClosed), sd) #t
sapply(select(test, nOpen, nClosed), sd) #n
```

#### Skew
```{r}
sapply(select(test, tOpen, tClosed), skew) #t
sapply(select(test, nOpen, nClosed), skew) #n
```
No skew

#### t.test
```{r}
#Time in Arms
t.test(test$tOpen, test$tClosed, paired = T)
```
SIGNIFICANTLY DIFFERENT!

```{r}
#Arm entries 
t.test(test$nOpen, test$nClosed, paired = T)
```

```{r}
lm(cbind(tOpen, tClosed) ~ Status, data = test)
```

